---
title: "coexpression_file"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install packages

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()
```

```{r}
install.packages(c("matrixStats", "Hmisc", "splines", "foreach", "doParallel", "fastcluster", "dynamicTreeCut", "survival"))
BiocManager::install()

```
```{r}
install.packages("BiocManager")
BiocManager::install(c("GO.db", "preprocessCore", "impute"))
```

```{r}
BiocManager::install("WGCNA")
```

#load data

```{r}
library(WGCNA);
options(stringsAsFactors = FALSE);
train_rnaseq = read.csv("rnaseq_geneset.csv", header = TRUE);
test_rnaseq = read.csv("test_rnaseq_geneset.csv", header = TRUE);
rownames(train_rnaseq) <- train_rnaseq$X
train_rnaseq$X <- NULL 
train_rnaseq <- t(train_rnaseq)
as.data.frame.matrix(train_rnaseq)
rownames(test_rnaseq) <- test_rnaseq$X
test_rnaseq$X <- NULL
test_rnaseq <- as.data.frame.matrix(test_rnaseq)
colnames(test_rnaseq) <- gsub("\\.", "-", colnames(test_rnaseq))

```
```{r}
colnames(train_rnaseq)
```

```{r}
colnames(test_rnaseq)

```


#check for too many missing values

```{r}
gsg = goodSamplesGenes(train_rnaseq, verbose = 3);
gsg$allOK
```

#check for outliers
```{r}
sampleTree = hclust(dist(train_rnaseq), method = "average");
# Plot the sample tree: Open a graphic output window of size 12 by 9 inches
# The user should change the dimensions if the window is too large or too small.
sizeGrWindow(8,4)
#pdf(file = "Plots/sampleClustering.pdf", width = 12, height = 9);
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,
cex.axis = 1.5, cex.main = 2)
```
#No significant outliners appear. 

#choose power

```{r}
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to=20, by=2))
# Call the network topology analysis function
sft = pickSoftThreshold(train_rnaseq, powerVector = powers, verbose = 5)
# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2));
cex1 = 0.9;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```
#Go for a soft power of 9.

#perform calculation
```{r}
net = blockwiseModules(train_rnaseq, power = 9,
TOMType = "unsigned", minModuleSize = 2,
reassignThreshold = 0, mergeCutHeight = 0.25,
numericLabels = TRUE, pamRespectsDendro = FALSE,
saveTOMs = TRUE,
saveTOMFileBase = "train_rna_geneset",
verbose = 3)
```

#check for number of modules (0 is for genes not in a module)
```{r}
table(net$colors)
```

```{r}
moduleLabels = net$colors
moduleColors = labels2colors(net$colors)
MEs = net$MEs;
save(MEs, moduleLabels, moduleColors,
file = "train_rnaseq_geneset-networkConstruction-auto.RData")
```

#ME values for genes that are in a module
```{r}
MEs$ME0 <- NULL
coexpression_train <- MEs
coexpression_train
```

#for train set:concat the expression data for genes outside modules with the calculated ME values for each module
#For modules' color label: turquoise = 1, grey = 0, blue = 2, brown = 3
```{r}
library(data.table)
library(dplyr)
mod0_name <- names(net$colors)[moduleColors=="grey"]
# train_no_moules <- subset(df, net$colors ==0)
# setDT(train_no_moules, keep.rownames = TRUE)[] #give a column name so can pull name out 
# train_genenomodules <- train_no_moules%>% pull(rn)
# train_genenomodules
final_train_coexpression <- train_rnaseq[,colnames(train_rnaseq) %in% mod0_name]
final_train_coexpression <- as.data.frame.matrix(final_train_coexpression)
final_train_coexpression <- merge(final_train_coexpression, coexpression_train, by="row.names")
rownames(final_train_coexpression) <- final_train_coexpression$Row.names
final_train_coexpression$Row.names <- NULL
final_train_coexpression
```

```{r}
write.csv(final_train_coexpression,"Final_train_coexpression.csv")
```



#get the name of genes within each module to pull the expression data out from test set

```{r}
mod1_name <- names(net$colors)[moduleColors=="turquoise"]
mod2_name <- names(net$colors)[moduleColors=="blue"]
mod3_name <- names(net$colors)[moduleColors=="brown"]
```

#form dataframes for each module of genes from test set
```{r}
df_1 <- t(test_rnaseq[,colnames(test_rnaseq) %in% mod1_name])
df_1 <- as.data.frame.matrix(df_1)

df_2 <- t(test_rnaseq[,colnames(test_rnaseq) %in% mod2_name])
df_2 <- as.data.frame.matrix(df_2)

df_3 <- t(test_rnaseq[,colnames(test_rnaseq) %in% mod3_name])
df_3 <- as.data.frame.matrix(df_3)
df_2

```

#Calculate ME for each module for the test set



```{r}
ei1 <- svd(df_1)$v
ei1 <- as.data.frame((ei1))
ei1 <-rowMeans(ei1)
ei1 <- as.data.frame((ei1))
row.names(ei1) <- row.names(test_rnaseq)
names(ei1)[1]<-"ME1"

ei2 <- svd(df_2)$v
ei2 <-rowMeans(ei2)
ei2 <- as.data.frame((ei2))
row.names(ei2) <- row.names(test_rnaseq)
names(ei2)[1]<-"ME2"


ei3 <- svd(df_3)$v
ei3 <- as.data.frame((ei3))
ei3 <-rowMeans(ei3)
ei3 <- as.data.frame((ei3))
row.names(ei3) <- row.names(test_rnaseq)
names(ei3)[1]<-"ME3"
```

#for test set: concat the expression data for genes outside modules with the calculated ME values for each module
```{r}
final_test_coexpression0 <- test_rnaseq[,colnames(test_rnaseq) %in% mod0_name]
final_test_coexpression1 <- merge(final_test_coexpression0,ei1, by="row.names")
rownames(final_test_coexpression1) <- final_test_coexpression1$Row.names
final_test_coexpression1$Row.names <- NULL
final_test_coexpression2 <- merge(final_test_coexpression1,ei2, by="row.names")
rownames(final_test_coexpression2) <- final_test_coexpression2$Row.names
final_test_coexpression2$Row.names <- NULL
final_test_coexpression <- merge(final_test_coexpression2,ei3, by="row.names")
rownames(final_test_coexpression) <- final_test_coexpression$Row.names
final_test_coexpression$Row.names <- NULL
final_test_coexpression

```

```{r}
write.csv(final_test_coexpression,"Final_test_coexpression.csv")
```




